Code for Breast Cancer Pain Study

### 2020-07-28 manuscript_code_20200722.Rmd
blight added

@@

unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
dmd$race.f = factor(dmd$race.f)

@@

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
# dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Vacancy",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR","Metastasis")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

@@

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(NA,NA)))
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(1,NA)))
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    tmp.sum[1,1] = 1
  }
  return(tmp.sum)
}

@@

compare_means(Pain ~ Occupation,data=final.data)
```

@@

x = xlist[2]
base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
p
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR","Metastasis")

explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR","Metastasis")

explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

@@

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.count)))) %>%
  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

>>>>

transform.p = function(x,digits=3) {
  y = as.character(round(x,digits=digits))
  y[which(y==0)] = "<.001"
  return(y)
}

@@

# unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
# dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
dmd$race.f[which(dmd$race.f %in% c("Other","Unknown"))] = "Unknown"
# cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
dmd$race.f = factor(dmd$race.f)

@@

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Vacancy",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    pval = compare_means(formula,data=final.data)
    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
  }
  return(tmp.sum)
}

@@

pval = compare_means(Pain ~ Occupation,data=final.data)
min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))])
```

@@

x = xlist[4]
base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
p
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")
